---
title: |
       | Extending \rlang with C++:
       | A Brief Introduction to Rcpp

# Use letters for affiliations
author:
  - name: Dirk Eddelbuettel
    affiliation: a
  - name: James Joseph Balamuta
    affiliation: b
address:
  - code: a
    address: Debian and R Projects; Chicago, IL, USA; \url{edd@debian.org}
  - code: b
    address: Depts of Informatics and Statistics, Univ. of Illinois at Urbana-Champaign; Champaign, IL, USA; \url{balamut2@illinois.edu}

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Eddelbuettel and Balamuta

# Place DOI URL or CRAN Package URL here
doi: "https://cran.r-project.org/package=Rcpp"

# Abstract
abstract: |
  \rlang has always provided an application programming interface (API) for extensions.
  Based on the \clang language, it uses a number of macros and other low-level constructs


# Acknowledgements
acknowledgements: |
  We thank Bob Rudis and Lionel Henry for excellent comments and suggestion on an earlier


# One or more keywords
keywords:
  - applications and case studies
  - statistical computing
  - computationally intensive methods
  - simulation

# Bibliography
bibliography: Rcpp

# Enable a watermark on the document
watermark: false

# Customize footer, eg by referencing the vignette
footer_contents: "Rcpp Vignette"

# Produce a pinp document
output: pinp::pinp

# Local additiona of a few definitions we use
header-includes: >
  \newcommand{\TODO}{\marginnote{TODO}}
  \newcommand{\rlang}{\textit{R }}
  \newcommand{\rlangns}{\textit{R}}
  \newcommand{\slang}{\textit{S }}
  \newcommand{\rcpp}{\textit{Rcpp }}
  \newcommand{\rcppns}{\textit{Rcpp}}
  \newcommand{\clang}{\textit{C }}
  \newcommand{\clangns}{\textit{C}}
  \newcommand{\cpp}{\textit{C++ }}
  \newcommand{\cppns}{\textit{C++}}
  \newcommand{\fortran}{\textit{Fortran }}
  \newcommand{\fortranns}{\textit{Fortran}}
  \newcommand{\python}{\textit{Python}}
  \newcommand{\julia}{\textit{Julia}}
  \newcommand{\pkg}[1]{\textbf{#1}}

vignette: >
  %\VignetteIndexEntry{Rcpp-introduction}
  %\VignetteKeywords{Rcpp, R, Cpp}
  %\VignettePackage{Rcpp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
library(Rcpp)
```


# Introduction

The \rlang language and environment \citep{R:main} has established itself as
both an increasingly dominant facility for data analysis, and the
*lingua franca* for statistical computing in both research and
application settings.


## Background



```{r evalCpp, eval=FALSE}
library("Rcpp")
evalCpp("2 + 2")
```


A graphical breakdown of all
aspects of a corresponding \cpp function is given in Figure \ref{fig:cpp-function-annotation}.

\begin{figure*}
  \begin{center}
    \includegraphics[width=5.5in]{figures/function_annotation_cpp.png}
    \caption{Graphical annotation of the \texttt{is\_odd\_cpp} function.}
    \label{fig:cpp-function-annotation}
  \end{center}
\end{figure*}

When using \rcppns, such \cpp functions can be directly embedded and compiled in an \rlang
script file through the use of the `cppFunction()` 



# Conclusion

\rlang has always provided mechanisms to extend it.  The bare-bones \clang API is already used

